Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-02-10 ribose-paper.RMD
updates to discussion

The main findings of the present study were that resistance training with glucose for five sessions did not affect muscle biology markers, measured as total RNA and -protein, rRNA and specific protein, compared to RT with placebo. Similarly, RT with glucose did not affect markers of skeletal muscle functionality such as muscular recovery, total session volume, and training intensity, compared to RT with placebo. There are studies suggesting that high glucose treatment impacts ribosome biogenesis in … cells (MARIAPPAN11; ZHAI12; TANAKA15), however, this effect does not seem to translate to human skeletal muscle cells in vivo.

To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle cells, making our hypotheses exploratory and comparisons not straight forwards. The current model of ribosome biogenesis suggests that changes in the accumulation of total RNA and rRNA early after the onset of an RT period predict long-term RT adaptations (FIGUEIREDO19; HAMMARSTROM20; HAMMARSTROM21).Therefore, we assumed 2 weeks to be a suitable time frame for a moderate RT volume to elicit changes in ribosome biogenesis, as similar observations have been made previously (FIGUEIREDO16; HAMMARSTROM20; HAMMARSTROM21). In Hammarström et al. (2020), untrained participants that increased ribosome biogenesis after two weeks (5 RT sessions) of RT, also responded well after the entire intervention (12 weeks, 31 sessions). Figueiredo and colleagues (2016) also observed markedly acute effects in RT-untrained individuals, where 47S pre-rRNA increased after a single RT session, passing baseline after 4hrs, peaking at 24hrs and remaining elevated until 48hrs post-exercise. The four mature rRNA did not measurably increase until after multiple sessions (FIGUEIREDO16). As such, increases in total RNA and rRNA observed in the present study are in line with previous observations (FIGUEIREDO16; HAMMARSTROM20; H21?). Further, the accumulation of total RNA and rRNA is also suggested to reach a plateau after approximately 5 RT-sessions, thus, in addition to the demanding RT- and dietary protocol and the exploratory nature of the hypotheses, we did not conduct a more long-term intervention. COMMENT ON PROTEIN DATA HERE?

As we did not observe any effect of ingesting glucose before and after RT compared to placebo in this design, there is little to suggest that glucose *per se* has any additional effect on ribosome biogenesis, as long as the daily energy intake is sufficient to support cellular energy levels required to drive this process. Looking at previous observations from *in vitro* studies of rodents, yeast, and non-muscle cell cultures (MARIAPPAN11; KIM13; TANAKA18; ZHAI12; HOPPE09; MURAYAMA08), one interesting notion would be how low-carb or glucose starvation would affect ribosome biogenesis in human skeletal muscle cells...

..Secondly, translational capacity is now regarded a vital factor in long-term RT-induced adaptations (IS IT?), as increased ribosome content provides a higher basal MPS, giving regulation of ribosome biogenesis a central role in muscle growth (FIGUEIREDO19). This is further substantiated by the association between blunted ribosome biogenesis and poor responses to RT (HAMMARSTROM20; STEC16), in addition to the lack of relationship between translational efficiency and long-term gains in skeletal muscle mass (FIGUEIREDO19). Hence, two weeks of moderate-volume unilateral RT should have been sufficient to elicit differences between exercise training with glucose and placebo. Supporting the notion that our intervention indeed was effective, we observed expected augmentation in accumulation of total RNA and expression of rRNA (CITE THIS?). As there were no differences in total RNA and rRNA between exercise training with glucose and placebo in the present study, it is therefore reasonable to suggest that there were no differences in ribosome biogenesis by proxy. Further, since glucose did not augment ribosome content after two weeks of RT compared to placebo, it is unlikely that glucose per se is beneficial to long-term adaptations in RT.

Indeed, our findings suggest that glucose ingestion per se before and after RT does not induce changes in the internal physiological milieu of human skeletal muscle cells with anabolic effects. It has been theorised that glucose may convey an anabolic signal through insulin, and thus operate as a signal mediator affecting upstream pathways of ribosome biogenesis (KUSNADI15; TANAKA18). Such an anabolic signal could then possibly increase PIC assembly, and consequently Pol I recruitment, initiating rDNA transcription and subsequent rRNA maturation, leading to increased ribosome content (FIGUEIREDO19; TANAKA18). INVESTIGATE THE INSULIN DATA AND COMMENT IT HERE. Another theory was that glucose ingestion would impact the cellular energy levels and thus, the energy-sensitive PIH1, mTORC1, ERK1/2, AMPK, and SIRT1 pathways (MARIAPPAN11; KIM13; TANAKA18; ZHAI12). The mammalian target of rapamycin, along with ERK1/2 and PIH1 up-regulates PIC assembly in high-energy conditions, while AMPK and possibly SIRT1 represses PIC assembly in low-energy conditions, based on in vitro rodent, yeast, and non-muscle cell cultures (MARIAPPAN11; KIM13; TANAKA18; ZHAI12; HOPPE09; MURAYAMA08). Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-IA and chromatin regulation suppression (KOS-BRAUN17; MARIAPPAN11; ZHAI12). In addition to UBF and SL-1, assembly of the PIC requires TIF-IA and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation (KUSNADI15). COMMENT ON PROTEIN ANALYSIS RESULTS HERE? As such, increasing energy supply and nutrient availability should in line with these theories potently stimulate PIC assembly and rDNA transcription. Along this notion, it does not seem apparent that the timing of glucose ingestion, for instance immediately before and after RT affects these processes, as glucose did not induce acute augmentation of ribosome biogenesis, despite the significant elevation of plasma glucose levels when exercising with glucose compared to placebo. Based on our observations, glucose per se has no acute effect on ribosome biogenesis, as the same effects can be observed with equal glucose ingestion 5-6hrs later.

Another possible explanation for the lack of effect of glucose compared to placebo, is that both treatments may have exercised with sufficient levels of nutrients and energy, meaning that mTORC1 had sufficient nutrient availability and, thus was not inhibited by AMPK due to energy deficiency (HOPPE09; KUSNADI15). Indeed, mTORC1 is characterised as energy-sensitive (DENNIS01), and therefore it might have been directly affected by high vs. low-glucose conditions (MARIAPPAN11). However, previous studies suggest that mTORC1 is not sensitive enough to register levels of ATP, rather depending on signals from AMPK to assess ATP levels (KIM10). This would arguably make AMPK the primary regulator of the potential impact of glucose on ribosome biogenesis, and thus, glucose has no direct effect on mTORC1 signalling (REFERENCES?). Importantly, intramuscular glycogen content was not measured in the present study, therefore we cannot determine whether cellular energy levels in fact were equal or not, or if there is reason to assume that AMPK-dependent mTORC1 signalling was different in the two treatments. Potentially, such data could give higher resolution to our data. Along with the robust increases in proxy markers of ribosome biogenesis, the equal macronutrient intake between treatments on pairwise consecutive days and plasma glucose levels not indicating energy deficiency during RT, potent mTORC1 stimulation without AMPK-dependent inhibition may be a plausible explanation. Together with previous studies on this topic (MARIAPPAN11; ZHAI12), there are currently scarce observations to suggest glucose carries a direct anabolic signal. As such, physiological levels of plasma glucose during RT, combined with a sufficient daily macronutrient intake, seem potent enough to supply the energy needed for ribosome biogenesis. The lacking differences between treatments in the biological data are further supported by our observations on functional data.

>>>>

The main findings of the present study were that resistance training with glucose for five sessions did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation associated proteins, compared to RT with placebo. Similarly, RT with glucose did not affect markers of skeletal muscle functionality such as muscular performance and -recovery, total session volume, and training intensity, compared to RT with placebo. There are studies suggesting that high vs. low glucose treatment/glucose starvation impacts ribosome biogenesis in yeast, rodent and human non-muscle cells (KIM13;MARIAPPAN11;TANAKA15;ZHAI12), however, this effect does not seem evident when total daily glucose intake is matched.

To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle cells, making our hypotheses exploratory and comparisons not straight forwards. The current model of ribosome biogenesis suggests that changes in the accumulation of total RNA and rRNA early after the onset of an RT period predict long-term RT adaptations (FIGUEIREDO19; HAMMARSTROM20; HAMMARSTROM21).Therefore, we assumed 2 weeks to be a suitable time frame for a moderate RT volume to elicit changes in ribosome biogenesis, as similar observations have been made previously (FIGUEIREDO16; HAMMARSTROM20; HAMMARSTROM21). In Hammarström et al. (2020), participants that increased ribosome biogenesis after two weeks (5 RT sessions) of RT, also responded well after the entire intervention (12 weeks, 31 sessions). Figueiredo and colleagues (2016) also observed markedly acute effects, where 47S pre-rRNA increased after a single RT session, passing baseline after 4hrs, peaking at 24hrs and remaining elevated until 48hrs post-exercise. The four mature rRNA did not measurably increase until after multiple sessions (FIGUEIREDO16). As such, increases in total RNA and rRNA observed in the present study are in line with previous observations (FIGUEIREDO16; HAMMARSTROM20; H21?), and also evident i moderately trained individuals, compared to the untrained or RT-untrained individuals of previous studies (FIGUEIREDO16;HAMMARSTROM20;H21?).

Further, the accumulation of total RNA and rRNA is also suggested to reach a plateau after approximately 5 RT-sessions, thus, in addition to the demanding RT- and dietary protocol and the exploratory nature of the hypotheses, we did not conduct a more long-term intervention. COMMENT ON PROTEIN DATA HERE?

As we did not observe any effect of ingesting glucose before and after RT compared to placebo in this design, there is little to suggest that glucose *per se* has any additional effect on ribosome biogenesis, as long as the daily energy intake is sufficient to support cellular energy levels required to drive this process. Looking at previous observations from *in vitro* studies of rodents, yeast, and non-muscle cell cultures (MARIAPPAN11; KIM13; TANAKA18; ZHAI12; HOPPE09; MURAYAMA08), one interesting notion would be how high glucose vs. low glucose or glucose starvation would affect ribosome biogenesis over time in human skeletal muscle cells. Time being of the essence in such a perspective, as previous *in vitro* studies have utilized cells with higher physiological rate than human skeletal muscle cells (KIM13;MARIAPPAN11;TANAKA15;ZHAI12).

Indeed, our findings suggest that glucose ingestion *per se* before and after RT does not induce changes in the internal physiological milieu of human skeletal muscle cells with anabolic effects. It has been theorised that glucose may convey an anabolic signal through insulin, and thus operate as a signal mediator affecting upstream pathways of ribosome biogenesis (KUSNADI15; TANAKA18). Such an anabolic signal could then possibly increase PIC assembly, and consequently Pol I recruitment, initiating rDNA transcription and subsequent rRNA maturation, leading to increased ribosome content (FIGUEIREDO19; TANAKA18). As significantly higher blood levels of c-peptide and glucose were observed in RT with glucose compared to placebo in the present study, without any difference in ribosome biogenesis, there is little to suggest that either insulin or glucose has any direct acute effects.

Another theory is that glucose ingestion leads to higher cellular energy levels compared to placebo, stimulating energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 in a dose-response relationship (KIM13;MARIAPPAN11;TANAKA15;ZHAI12). The mTORC1, along with ERK1/2 and PIH1 up-regulates PIC assembly during high-energy conditions, while AMPK and possibly SIRT1 represses PIC assembly during low-energy conditions, as observed in rodent (KILDE), yeast (KILDE), and non-muscle cells (KILDE) *in vitro*. Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-IA and chromatin regulation suppression (KOS-BRAUN17; MARIAPPAN11; ZHAI12). In addition to UBF and SL-1, assembly of the PIC requires TIF-IA and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation (KUSNADI15). As such, increasing energy supply and nutrient availability should in line with these theories potently stimulate PIC assembly and rDNA transcription. Importantly, these pathways were not directly investigated in the present study. However, downstream targets (UBF, c-Myc, rpS6?) remained unaffected by glucose compared to placebo ingestion, suggesting there is not a dose-response relationship between muscle cellular energy levels and ribosome biogenesis during an RT-intervention (Not sure if I can say this).

A weakness in the present study, is the lack of data on intracellular glucose levels in the muscle cells. Thus, it cannot be said for sure whether there were differences between glucose and placebo or not.

A possible explanation for the lack of effect of glucose compared to placebo, is that both treatments may have exercised with sufficient levels of nutrients and energy, meaning that mTORC1 had sufficient nutrient availability and, thus was not inhibited by AMPK due to energy deficiency (HOPPE09; KUSNADI15). Indeed, mTORC1 is characterised as energy-sensitive (DENNIS01), and therefore it might have been directly affected by high vs. low-glucose conditions (MARIAPPAN11). However, previous studies suggest that mTORC1 is not sensitive enough to register levels of ATP, rather depending on signals from AMPK to assess ATP levels (KIM10). This would arguably make AMPK the primary regulator of the potential impact of glucose on ribosome biogenesis, and thus, glucose has no direct effect on mTORC1 signalling (REFERENCES?). Importantly, intramuscular glycogen content was not measured in the present study, therefore we cannot determine whether cellular energy levels in fact were equal or not, or if there is reason to assume that AMPK-dependent mTORC1 signalling was different in the two treatments. Potentially, such data could give higher resolution to our data. Along with the robust increases in proxy markers of ribosome biogenesis, the equal macronutrient intake between treatments on pairwise consecutive days and plasma glucose levels not indicating energy deficiency during RT, potent mTORC1 stimulation without AMPK-dependent inhibition may be a plausible explanation. Together with previous studies on this topic (MARIAPPAN11; ZHAI12), there are currently scarce observations to suggest glucose carries a direct anabolic signal. As such, physiological levels of plasma glucose during RT, combined with a sufficient daily macronutrient intake, seem potent enough to supply the energy needed for ribosome biogenesis. The lacking differences between treatments in the biological data are further supported by our observations on functional data.
